---
title: "DATA 607 - Assignment2 [R and SQL]"
author: "Simon63"
date: "February 11, 2018"
output: html_document
params:
  dbPwd: enter password
---
***
Using Knit with Parameters...  
for DB password  
**params:  dbPwd:** enter password  

***
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Package Installs (optional)
install.packages("RMySQL");  
install.packages("ggplot2");

### Libraries
```{r libraries, warning=FALSE, message=FALSE}
library(RCurl);
library(DBI);
library(RMySQL);
library(ggplot2);
```

### Getting MySQL Driver and Connection
```{r mysql_con}
dbDrv <- dbDriver("MySQL");
con <- dbConnect(dbDrv, user="root", password=params$dbPwd, host="localhost", port=3306, dbname="movies");
```

### Using DBI Functions
```{r dbi_func}
dbListTables(con);
dbReadTable(con, "movie");
sql <- "SELECT 
     p.name, m.title, IFNULL(r.rank, 'NA') AS rank
 FROM
 person p
 LEFT JOIN
 rating r ON p.id = r.person_id
 INNER JOIN
 movie m ON r.movie_id = m.id;";
res <- dbGetQuery(con, sql);
knitr::kable(
  head(res),
  caption = "dbGetQuery result set [head]"
)
```

###Using {sql} markdown to execute SQL
```{sql sql_rmd, connection=con, max.print = 10, tab.cap="Same result as dbGetQuery [showing first 10 rows]"}
SELECT 
    p.name, m.title, IFNULL(r.rank, 'NA') AS rank
FROM
    person p
        LEFT JOIN
    rating r ON p.id = r.person_id
        INNER JOIN
    movie m ON r.movie_id = m.id;
```

###Plotting a Bar chart
Note: SQL output.var="res"
```{sql sql_chart, connection=con, output.var="res"}
SELECT 
    m.id movie_id, AVG(r.rank) AS avg_rank
FROM
    rating r 
        INNER JOIN
    movie m ON r.movie_id = m.id
GROUP BY m.title
ORDER BY AVG(r.rank) DESC;
```

```{r chart}
barplot(height=res$avg_rank, names.arg = res$movie_id, xlab = "Movie ID", main = "Movie Ratings")
```
```{sql sql_chart_legend, connection=con, echo=FALSE, tab.cap="X Axis Legend"}
SELECT 
    m.id AS "Movie ID", m.title "Movie Title"
FROM
    movie m
ORDER BY m.id;
```

####Close DB Connection
```{r db_close}
dbDisconnect(con)
```

### Using {sql, eval=FALSE} markdown to only display SQL (with syntax hightlighting)
Below is the scipt for setting up "movies" database
```{sql sql_display, eval=FALSE}
/* These statements should be executed only once */
CREATE DATABASE movies;
USE movies;
/*************************************************/

DROP TABLE IF EXISTS person;
DROP TABLE IF EXISTS movie;
DROP TABLE IF EXISTS rating;

CREATE TABLE person (
    id INTEGER NOT NULL PRIMARY KEY,
    name VARCHAR(30) NOT NULL UNIQUE
);

LOAD DATA INFILE 'C:/ProgramData/MySQL/MySQL Server 5.7/Uploads/movies/person.csv' 
INTO TABLE person
FIELDS TERMINATED BY ',' 
OPTIONALLY ENCLOSED BY '"'
LINES TERMINATED BY '\r\n'
IGNORE 1 ROWS;

CREATE TABLE movie (
    id INTEGER NOT NULL PRIMARY KEY,
    title VARCHAR(60) NOT NULL UNIQUE KEY
);

LOAD DATA INFILE 'C:/ProgramData/MySQL/MySQL Server 5.7/Uploads/movies/movie.csv' 
INTO TABLE movie
FIELDS TERMINATED BY ',' 
OPTIONALLY ENCLOSED BY '"'
LINES TERMINATED BY '\r\n'
IGNORE 1 ROWS;

CREATE TABLE rating (
    person_id INTEGER NOT NULL REFERENCES person (id),
    movie_id INTEGER NOT NULL REFERENCES movie (id),
    rank INTEGER COMMENT 'rating score of each movie by a person on a scale of 1 to 5',
    CHECK (IFNULL(rank, 1) BETWEEN 1 AND 5),
    CONSTRAINT PRIMARY KEY (person_id , movie_id),
    INDEX (person_id , movie_id) COMMENT 'good practice to define an index with first column being a reference to another table (FK)',
    INDEX (movie_id , person_id) COMMENT 'index on a foreign key (FK) column'
);

LOAD DATA INFILE 'C:/ProgramData/MySQL/MySQL Server 5.7/Uploads/movies/rating.csv' 
INTO TABLE rating
FIELDS TERMINATED BY ',' 
OPTIONALLY ENCLOSED BY '"'
LINES TERMINATED BY '\r\n'
IGNORE 1 ROWS
(person_id, movie_id, @rank)
SET rank = nullif(@rank, '');
```

